## Practical Machine Learning Assignment


-Describe how model was built (machine learning algorithm to predict activity quality from activity monitors)
-Describe how cross-validation was used
-Estimate out of sample error using cross-validation
-Why I made the choices I did
-R markdown file
-HTML file < 2,000 words; < 5 figures
-Send link to Github repo with a gh-pages branch



```{r}
library(caret)
library(randomForest)
```


Data source:  http://groupware.les.inf.puc-rio.br/har; 2014-11-14

```{r}
trainingURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

if(!file.exists("pml-training.csv")){download.file(url=trainingURL,destfile="pml-training.csv")}
if(!file.exists("pml-testing.csv")){download.file(url=testingURL,destfile="pml-testing.csv")}

training <- read.csv(file="pml-training.csv",header=TRUE,stringsAsFactors=FALSE)
testing <- read.csv(file="pml-testing.csv",header=TRUE,stringsAsFactors=FALSE)
```



```{r}
training$classe <- as.factor(training$classe)

atts <- c(7:11,37:49,60:68,84:86,113:124,140,151:159)

for(i in atts){
  training[,i] <- as.numeric(training[,i])
  testing[,i] <- as.numeric(testing[,i])
}
```



```{r}
inTrain <- createDataPartition(y=training$classe,p=0.7,list=FALSE)
train <- training[inTrain,]
test <- training[-inTrain,]
```


don't actually run the model since it takes too long:

```{r eval=FALSE}
modFit <- train(classe~.,method="rf",data=train[,c(atts,160)])
```


rest of the code:

```{r}
modFit
modFit$finalModel
testConf <- table(test$classe,predict(modFit,test))
sum(diag(testConf)) / sum(testConf)
```


Plot(s):

```{r, echo=FALSE}

```

